---
title: "Final_model"
author: "Jakub"
date: "25 října 2016"
output: html_document
---

```{r param, include=FALSE}
library(deSolve)

options("scipen" = 10)
N <- 1
a <- 10000 #time span
m <- 1 #time step

pbeta_w <- 0.04 #transmission parameter of Iw
pbeta_z <- 0.015 #transmission parameter of Iz
pbeta_z2 <- 0.04 #transmission parameter of Iz when there is no transmission cost
pbeta_z_test <- 0.015 #transmission parameter of Iz when there is no transmission cost



pgammaa <- 0.01 #innate recovery rate
pgamma_tilt <- 0.1 #recovery rate of Iw when treated by antibiotics


prho <- 0.05 #probability of clinical failure
prho_test <- 0.05

peps <- 0.10 #initial value of the ratio of treated patients
peps_test <- 0.10 #initial value of the ratio of treated patients





prhoi <- seq(0,0.10,by=0.001)
prhoi_ext <- seq(0,0.50,by=0.001)
pepsi <- seq(0,1,by=0.01)


prop <- 0.01 



conv_crit <- 0.00001


```



\begin{aligned}

\frac{dS}{dt} &=- \beta_w I_wS - \beta_z I_z S + \gamma  I_z + \gamma(1 - \epsilon)I_w + \gamma `  \epsilon  I_w \\

\frac{dI_w}{dt} &=\beta_w I_wS + (1-\epsilon)\beta_wI_wI_z - \gamma(1 - \epsilon)I_w - \gamma `  \epsilon  I_w - \epsilon\beta_zI_zI_w - \rho\epsilon I_w \\

\frac{dI_z}{dt} &=\beta_z I_zS + \epsilon\beta_zI_zI_w + \rho\epsilon I_w - \gamma  I_z - (1-\epsilon)\beta_wI_wI_z \\


\end{aligned}
 
```{r trial, include=FALSE}
parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z_test, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps_test, rho=prho_test)

state <- c(S=0.1, Iw=0.1, Iz=0.8)

times <- seq(0, a, by=m)

SIS=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*Iw*S - beta_z*Iz*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*Iw*S + (1-eps)*beta_w*Iw*Iz - eps*beta_z*Iz*Iw - rho*eps*Iw - gamma*(1-eps)*Iw - gamma_prim*eps*Iw
        dIz = beta_z*Iz*S + eps*beta_z*Iz*Iw + rho*eps*Iw - (1-eps)*beta_w*Iw*Iz - gamma*Iz
        
        list(c(dS, dIw, dIz))
       })
}

out=as.data.frame(
  ode(y=state, times=times, func=SIS, parms=parameters)
)

```


```{r main_trial, echo=FALSE, fig.align='center'}

plot(times, out$S, main="Trial", xlab="time (steps)", ylab="ratio", ylim=c(0,1), lty=1, lwd=2, type="l")
lines(times, out$Iw, lty=3, lwd=2)
lines(times, out$Iz, lty=2, lwd=2)

legend("topright" , inset=.05, c("Iw", "Iz", "S" ), lty=c(3,2,1), lwd=c(2,2,2))
```



<h2> $\varepsilon$ variation </h2>

``` {r steady_state_eps, include=FALSE, cache=TRUE}

parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps, rho=prho)

state <- c(S=1-2*prop, Iw=prop, Iz=prop)

times <- seq(0, a, by=m)


root_main=function(times, state, parameters){
  dstate <- unlist(SIS(times, state, parameters))
  sum(abs(dstate)) - conv_crit
  }

exploit_param_eps=function(x) {
  parameters[5] <- x
  out <- lsodar(func=SIS, y=state, parms=parameters, times=times, rootfunc=root_main)
  return(tail(out, n=1))
  }

steady_eps <- do.call(rbind, lapply(pepsi, exploit_param_eps))
steady_eps <- unlist(steady_eps)

steady_time_eps <- steady_eps[,1]
steady_S_eps <- steady_eps[,2]
steady_Iw_eps <- steady_eps[,3]
steady_Iz_eps <- steady_eps[,4]
steady_QA_Iw_eps <- steady_Iw_eps*pepsi
steady_QA_Iw_Iz_eps <- (steady_Iw_eps+steady_Iz_eps)*pepsi

```

```{r steady_state_eps_plot, echo=FALSE, fig.align='center'}

plot(pepsi, steady_Iw_eps, main="Steady-states : treatment variations", xlab=expression(epsilon), ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(pepsi, steady_Iz_eps, lty=2, lwd=2)
lines(pepsi, steady_S_eps, lty=1, lwd=2)

legend("topleft", inset=.05, c("S", "Iz", "Iw"), lty=c(1,2,3), lwd=rep(2,3))


#zoom : control for no interhost compet

pepsi2 <- seq(0,0.20,by=0.01)
steady_Iw_eps2 <- steady_Iw_eps[1:length(pepsi2)]
steady_Iz_eps2 <- steady_Iz_eps[1:length(pepsi2)]
steady_S_eps2 <- steady_S_eps[1:length(pepsi2)]

plot(pepsi2, steady_Iw_eps2, main="Steady-states : treatment variations", xlab=expression(epsilon), ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(pepsi2, steady_Iz_eps2, lty=2, lwd=2)
lines(pepsi2, steady_S_eps2, lty=1, lwd=2)

legend("topleft", inset=.05, c("S", "Iz", "Iw"), lty=c(1,2,3), lwd=rep(2,3))

#zoom : control for no sec resistance partial model

pepsi3 <- seq(0,0.30,by=0.01)
steady_Iw_eps3 <- steady_Iw_eps[1:length(pepsi3)]
steady_Iz_eps3 <- steady_Iz_eps[1:length(pepsi3)]
steady_S_eps3 <- steady_S_eps[1:length(pepsi3)]

plot(pepsi3, steady_Iw_eps3, main="Steady-states : treatment variations", xlab=expression(epsilon), ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(pepsi3, steady_Iz_eps3, lty=2, lwd=2)
lines(pepsi3, steady_S_eps3, lty=1, lwd=2)

legend("topleft", inset=.05, c("S", "Iz", "Iw"), lty=c(1,2,3), lwd=rep(2,3))





plot(pepsi, steady_Iw_eps, main="Steady-states : treatment variations", xlab=expression(epsilon), ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(pepsi, steady_Iz_eps, lty=2, lwd=2)
lines(pepsi, steady_S_eps, lty=1, lwd=2)
lines(pepsi, steady_QA_Iw_eps, lty=1, lwd=2, col='red')
lines(pepsi, steady_QA_Iw_Iz_eps, lty=2, lwd=2, col='red')

legend("topright" , inset=.05, c("Iw", "Iz", "S", 'QA : Iw treated', 'QA : all treated' ), lty=c(3,2,1,1,2), lwd=c(2,2,2,2,2), col=c(rep('black',3), rep('red',2)))

```


<h2> Introduction of QA </h2>


```{r steady_state_QA_plot, echo=FALSE, fig.align='center'}

plot(pepsi, steady_QA_Iw_eps, main="QA : treatment variations", xlab=expression(epsilon), ylab="QA", ylim=c(0,0.4), lty=1, lwd=2, type="l", col='red')


lines(pepsi, steady_QA_Iw_Iz_eps, lty=2, lwd=2, col='red')

legend("topright" , inset=.05, c('QA : Iw treated', 'QA : all treated' ), lty=c(1,2), lwd=c(2,2), col=c(rep('red',2)))

```


<h2> $\rho$ variation </h2>

``` {r steady_state_rho, include=FALSE, cache=TRUE}

parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps, rho=prho)

state <- c(S=1-2*prop, Iw=prop, Iz=prop)

times <- seq(0, a, by=m)


root_main=function(times, state, parameters){
  dstate <- unlist(SIS(times, state, parameters))
  sum(abs(dstate)) - conv_crit
  }

exploit_param_rho=function(x) {
  parameters[6] <- x
  out <- lsodar(func=SIS, y=state, parms=parameters, times=times, rootfunc=root_main)
  return(tail(out, n=1))
  }

steady_rho <- do.call(rbind, lapply(prhoi, exploit_param_rho))
steady_rho <- unlist(steady_rho)

steady_time_rho <- steady_rho[,1]
steady_S_rho <- steady_rho[,2]
steady_Iw_rho <- steady_rho[,3]
steady_Iz_rho <- steady_rho[,4]

steady_rho_ext <- do.call(rbind, lapply(prhoi_ext, exploit_param_rho))
steady_rho_ext <- unlist(steady_rho_ext)

steady_time_rho_ext <- steady_rho_ext[,1]
steady_S_rho_ext <- steady_rho_ext[,2]
steady_Iw_rho_ext <- steady_rho_ext[,3]
steady_Iz_rho_ext <- steady_rho_ext[,4]

```

```{r steady_state_rho_plot, echo=FALSE, fig.align='center'}
plot(prhoi, steady_Iw_rho, main="Steady-states : treatment failure variations", xlab=expression(rho), ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(prhoi, steady_Iz_rho, lty=2, lwd=2)
lines(prhoi, steady_S_rho, lty=1, lwd=2)

legend("topright" , inset=.05, c("Iw", "Iz", "S" ), lty=c(3,2,1), lwd=c(2,2,2))

```

```{r steady_state_rho_ext_plot, echo=FALSE, fig.align='center'}
plot(prhoi_ext, steady_Iw_rho_ext, main="Steady-states : treatment failure variations", xlab=expression(rho), ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(prhoi_ext, steady_Iz_rho_ext, lty=2, lwd=2)
lines(prhoi_ext, steady_S_rho_ext, lty=1, lwd=2)

legend("topright" , inset=.05, c("Iw", "Iz", "S" ), lty=c(3,2,1), lwd=c(2,2,2))

```

<h3> $\rho$ variation without transmission cost for resistant strain</h3>


``` {r steady_state_no_cost_rho, include=FALSE, cache=TRUE}
prhoi <- seq(0,0.10,by=0.001)

parameters2 <- c(beta_w=pbeta_w, beta_z=pbeta_z2, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps, rho=prho)

state2 <- c(S=1-2*prop, Iw=prop, Iz=prop)

times2 <- seq(0, a, by=m)


root_main2=function(times2, state2, parameters2){
  dstate <- unlist(SIS(times2, state2, parameters2))
  sum(abs(dstate)) - conv_crit
  }

exploit_param_rho2=function(x) {
  parameters2[6] <- x
  out <- lsodar(func=SIS, y=state2, parms=parameters2, times=times2, rootfunc=root_main2)
  return(tail(out, n=1))
  }

steady_rho2 <- do.call(rbind, lapply(prhoi, exploit_param_rho2))
steady_rho2 <- unlist(steady_rho2)

steady_time_rho2 <- steady_rho2[,1]
steady_S_rho2 <- steady_rho2[,2]
steady_Iw_rho2 <- steady_rho2[,3]
steady_Iz_rho2 <- steady_rho2[,4]

steady_rho2_ext <- do.call(rbind, lapply(prhoi_ext, exploit_param_rho2))
steady_rho2_ext <- unlist(steady_rho2_ext)

steady_time_rho2_ext <- steady_rho2_ext[,1]
steady_S_rho2_ext <- steady_rho2_ext[,2]
steady_Iw_rho2_ext <- steady_rho2_ext[,3]
steady_Iz_rho2_ext <- steady_rho2_ext[,4]

```

```{r steady_state_rho_no_cost_plot, echo=FALSE, fig.align='center'}
plot(prhoi, steady_Iw_rho2, main="Steady-states : treatment failure variations \n when without transmission cost", xlab=expression(rho), ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(prhoi, steady_Iz_rho2, lty=2, lwd=2)
lines(prhoi, steady_S_rho2, lty=1, lwd=2)

legend("topright" , inset=.05, c("Iw", "Iz", "S" ), lty=c(3,2,1), lwd=c(2,2,2))

```

```{r steady_state_rho_ext_no_cost_plot, echo=FALSE, fig.align='center'}
plot(prhoi_ext, steady_Iw_rho2_ext, main="Steady-states : treatment failure variations \n when without transmission cost", xlab=expression(rho), ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(prhoi_ext, steady_Iz_rho2_ext, lty=2, lwd=2)
lines(prhoi_ext, steady_S_rho2_ext, lty=1, lwd=2)

legend("topright" , inset=.05, c("Iw", "Iz", "S" ), lty=c(3,2,1), lwd=c(2,2,2))

```

```{r average_stay_within_Iw, echo=FALSE, fig.align='center'}
epsico <- seq(0, 0.18, by=0.01)
epsico_ext <- seq(0, 1, by=0.01)

average_Iw <- 1/((1-epsico)*pgammaa + epsico*pgamma_tilt)
average_Iw_ext <- 1/((1-epsico_ext)*pgammaa + epsico_ext*pgamma_tilt)

plot(epsico, average_Iw, main="Average duration of sensitive infection", xlab=expression(epsilon), ylab="time (steps)", lty=1, lwd=2, type="l", col="green", ylim=c(0,100))

plot(epsico_ext, average_Iw_ext, main="Average duration of sensitive infection", xlab=expression(epsilon), ylab="time (steps)", lty=1, lwd=2, type="l", col="green", ylim=c(0,100))

```