---
title: "Supermodel"
author: "Jakub"
date: "18 října 2016"
output: html_document
---

```{r include=FALSE}
library(deSolve)
options("scipen" = 10)
N <- 1
a <- 2000
pbeta_w <- 0.04
pbeta_z <- 0.015
pgammaa <- 0.01
pgamma_tilt <- 0.1
pq <- 0.5
prho <- 0.5
pphi <- 0.05
peps <- 0.15
m <- 1
prop <- 0.01

# choice of model
pd <- 0   #coexistence in single host
pr <- 0   #replacement of Iz by Iw via interhost infection
pu <- 1   #Iw to Iz conversion
pb <- 1   #Iz to Iw conversion


```



<center> <h1> Synthesis of models mentionned in Spicknall`s review </h1> </center>

\begin{aligned}

\frac{dS}{dt} &=- {\beta_w} \times({I_w} + {d}\times{q}\times{I_{wz}})\times{S} - {\beta_z} \times({I_z} + {d}\times{q}\times{I_{wz}})\times{S} + {\gamma} \times {I_z} + {\gamma}\times(1 - {\epsilon})\times{I_w} + {\gamma `} \times {\epsilon} \times {I_w} \\

\frac{dI_w}{dt} &={\beta_w} \times({I_w} + {d}\times{q}\times{I_{wz}})\times{S} + {r}\times({\beta_w} - {u}\times{\rho}\times{\epsilon}\times{I_w} - {\beta_z})\times{I_w}\times{I_z} + {d}\times{\gamma} \times {I_{wz}} - {d}\times{\beta_z} \times({I_z} + {q}\times{I_{wz}})\times{I_w} - {\gamma}\times(1 - {\epsilon})\times{I_w} - {\gamma `}\times {\epsilon} \times {I_w} -  {b}\times{\phi}\times(1-{\epsilon})\times{I_z}\\

\frac{dI_z}{dt} &={\beta_z} \times({I_z} + {d}\times{q}\times{I_{wz}})\times{S} + {u}\times{\rho}\times{\epsilon}\times{I_w} - {r}\times({\beta_w} - {\beta_z})\times{I_w}\times{I_z} + {d}\times{\gamma}\times(1 - {\epsilon})\times{I_{wz}} + {d}\times{\gamma `} \times {\epsilon} \times {I_{wz}} - {\beta_w} \times({I_w}+{d}\times{q}\times{I_{wz}})\times{I_z} - {\gamma} \times {I_z} + {b}\times{\phi}\times(1-{\epsilon})\times{I_z}\\

\frac{dI_{wz}}{dt} &={d}\times({\beta_z} \times({I_z} + {q}\times{I_{wz}})\times{I_w} + {\beta_w} \times({I_w} + {q}\times{I_{wz}})\times{I_z} - {\gamma} \times {I_{wz}} -{\gamma}\times(1 - {\epsilon})\times{I_{wz}} - {\gamma `} \times {\epsilon} \times {I_{wz}}) \\

 \end{aligned}
 
 
 
 
```{r supermodel_def, include=FALSE}
parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps, q=pq, rho=prho, phi=pphi, d=pd, r=pr, u=pu, b=pb )

state <- c(S=1-2*prop, Iw=prop, Iz=prop, Iwz=0)

times=seq(0, a, by=m)

SIS_supermodel=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*(Iw+d*q*Iwz)*S - beta_z*(Iz+d*q*Iwz)*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*(Iw+d*q*Iwz)*S + d*gamma*Iwz - d*beta_z*(Iz+q*Iwz)*Iw - gamma*(1-eps)*Iw - gamma_prim*eps*Iw + r*(beta_w-beta_z)*Iw*Iz - u*rho*eps*Iw + b*phi*(1-eps)*Iz
        dIz = beta_z*(Iz+d*q*Iwz)*S + d*gamma*(1-eps)*Iwz + gamma_prim*eps*Iwz - d*beta_w*(Iw+q*Iwz)*Iz - gamma*Iz - r*(beta_w-beta_z)*Iw*Iz + u*rho*eps*Iw - b*phi*(1-eps)*Iz
        dIwz = d*(beta_w*(Iw+q*Iwz)*Iz + beta_z*(Iz+q*Iwz)*Iw - gamma*Iwz - gamma*(1-eps)*Iwz - gamma_prim*eps*Iwz)
        list(c(dS, dIw, dIz, dIwz))
       })
}


```


```{r super_trial, echo=FALSE, fig.align='center'}
require(deSolve)

out_super=as.data.frame(
  ode(y=state, times=times, func=SIS_supermodel, parms=parameters)
)

plot(times, out_super$S, main="Trial : bidirectional", xlab="time (steps)", ylab="ratio", ylim=c(0,1), lty=1, lwd=2, type="l")
lines(times, out_super$Iw, lty=3, lwd=2)
lines(times, out_super$Iz, lty=2, lwd=2)
lines(times, out_super$Iwz, lty=4, lwd=2)

legend("topleft" , inset=.05, c("S Susceptible","Iw Sensitive strain", "Iz Resistant strain", "Iwz Double infected" ), lty=c(1,3,2,4), lwd=c(2,2,2,2))
```
