---
title: "Spicknall"
author: "Jakub"
date: "4th October"
output: html_document
---



```{r include=FALSE}
library(deSolve)
options("scipen" = 10)
N=10000
a=50
pbeta_w=0.0001
pbeta_z=0.00005
pgammaa=0.005
pgamma_tilt=0.5
pq=0.25
prho=0.1
pphi=0.01
peps=0.7

```

<center> <h1> Models based on Spicknall`s review </h1> </center> 

The models include the susceptible population S, the population infected by sensitive strain I~w~, the population infected by resistant strain I~z~ and eventually the population infected by both I~zw~. Every model starts with S=`r N-2`, I~w~=1 and I~z~=1 at t=0. We then follow the evolution of these population monthly during `r a` years.

The mechanism of resistance has its cost in terms of transmission (for example because of lower reproduction rate of bacteria and therefore less bacteria are transmitted during the contact). This phenomenon is taken into account by the choice of  $\beta$~z~ and $\beta$~w~. $\beta$~z~ is lower than $\beta$~w~ and therefore the transmission takes place at a lower rate in resistant strain.

<center> <h2> Single-strain model </h2> </center>

This model is based on an independent evolution of the resistant and sensitive strain and doesn?t therefore incude any competition.

The model is defined by following equations:


 \begin{aligned}
 \frac{dS_1}{dt} &= -\beta_w \times{I_w}\times{S_1} +  \gamma\times(1-\epsilon)\times{I_w} + \gamma ` \times \epsilon \times {I_w} \\
 
 \frac{dI_w}{dt} &=\beta_w \times{I_w}\times{S_1} - \gamma\times(1-\epsilon)\times{I_w}- \gamma ` \times \epsilon \times {I_w} \\
 
 \frac{dS_2}{dt} &= -\beta_z \times{I_z}\times{S_2} + \gamma \times {I_z}  \\
 
 \frac{dI_z}{dt} &=\beta_z \times{I_z}\times{S_2} - \gamma \times {I_z} \\
 
 \end{aligned}
 





```{r single_strain_def, include=FALSE}
parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps )

state <- c(S1=N-1, S2=N-1, Iw=1, Iz=1)

times=seq(0, a, by=1/12)

SIS_single_strain=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS1 = - beta_w*Iw*S1 + gamma*(1-eps)*Iw + gamma_prim*eps*Iw
        dS2 = - beta_z*Iz*S2 + gamma*Iz
        dIw = beta_w*Iw*S1 - gamma*(1-eps)*Iw - gamma_prim*eps*Iw
        dIz = beta_z*Iz*S2 - gamma*Iz
        list(c(dS1, dS2, dIw, dIz))
       })
}

require(deSolve)

out=as.data.frame(
  ode(y=state, times=times, func=SIS_single_strain, parms=parameters)
)



```




```{r single_strain_trial, echo=FALSE, fig.align='center'}
plot(times, out$Iw/N, main="Plot n°1 - Single-strain model", xlab="time (yrs)", ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(times, out$Iz/N, lty=2, lwd=2)

double_inf <- out$Iw*out$Iz

lines(times, double_inf/N^2, lty=4, lwd=2)

legend("topleft" , inset=.05, c("Iw Sensitive strain", "Iz Resistant strain", "Iwz Double infected" ), lty=c(3,2,4), lwd=c(2,2,2))
```



__Plot n°1 Single-strain model__: $\beta$ ~w~ = `r pbeta_w`, $\beta$ ~z~ = `r pbeta_z`, $\gamma$ = `r pgammaa`, $\gamma$` = `r pgamma_tilt` and $\epsilon$ = `r peps` are the chosen values of parameters. The $\delta$t = 1/12 year i.e. 1 month and we follow the evolution of population infected with sensitive strain (_dotted line_), with resistant strain (_dashed line_) and with both (_dashed-dotted line_) over `r a` years.


<center> <h2> Superinfection model </h2> </center>

In this model, the hosts can be infected with both resistant and sensitive strains, I~wz~ corresponding to this subpopulation. This superinfected subpopulation of host can then transmit the one or another strain but with dampened contagiousness, that is represented by the ratio q in the model.

The model is defined by following equations:


\begin{aligned}

\frac{dS}{dt} &=-{\beta_w} \times({I_w}+{q}\times{I_{wz}})\times{S} -{\beta_z} \times({I_z}+{q}\times{I_{wz}})\times{S} + {\gamma} \times {I_z} + {\gamma}\times(1-{\epsilon})\times{I_w} + {\gamma `} \times {\epsilon} \times {I_w} \\

\frac{dI_w}{dt} &={\beta_w} \times({I_w}+{q}\times{I_{wz}})\times{S} + {\gamma} \times {I_{wz}}  - {\beta_z} \times({I_z}+{q}\times{I_{wz}})\times{I_w}- {\gamma}\times(1-{\epsilon})\times{I_w}- {\gamma `}\times {\epsilon} \times {I_w} \\

\frac{dI_z}{dt} &={\beta_z} \times({I_z}+{q}\times{I_{wz}})\times{S} + {\gamma}\times(1-{\epsilon})\times{I_{wz}} + {\gamma `} \times {\epsilon} \times {I_{wz}} - {\beta_w} \times({I_w}+{q}\times{I_{wz}})\times{I_z} - {\gamma} \times {I_z} \\

\frac{dI_{wz}}{dt} &={\beta_z} \times({I_z}+{q}\times{I_{wz}})\times{I_w}+{\beta_w} \times({I_w}+{q}\times{I_{wz}})\times{I_z}-{\gamma} \times {I_{wz}} -{\gamma}\times(1-{\epsilon})\times{I_{wz}} - {\gamma `} \times {\epsilon} \times {I_{wz}} \\

 \end{aligned}







```{r superinfection_def, include=FALSE}
parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps, q=pq )

state <- c(S=N-2, Iw=1, Iz=1, Iwz=0)

times=seq(0, a, by=1/12)

SIS_superinf=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*(Iw+q*Iwz)*S - beta_z*(Iz+q*Iwz)*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*(Iw+q*Iwz)*S + gamma*Iwz - beta_z*(Iz+q*Iwz)*Iw - gamma*(1-eps)*Iw - gamma_prim*eps*Iw 
        dIz = beta_z*(Iz+q*Iwz)*S + gamma*(1-eps)*Iwz + gamma_prim*eps*Iwz - beta_w*(Iw+q*Iwz)*Iz - gamma*Iz 
        dIwz = beta_w*(Iw+q*Iwz)*Iz + beta_z*(Iz+q*Iwz)*Iw - gamma*Iwz - gamma*(1-eps)*Iwz - gamma_prim*eps*Iwz
        list(c(dS, dIw, dIz, dIwz))
       })
}

require(deSolve)

out_super=as.data.frame(
  ode(y=state, times=times, func=SIS_superinf, parms=parameters)
)

```




```{r superinfection_trial, echo=FALSE, fig.align='center'}
plot(times, out_super$S/N, main="Plot n°2 - Superinfection model", xlab="time (yrs)", ylab="ratio", ylim=c(0,1), lty=1, lwd=2, type="l")
lines(times, out_super$Iw/N, lty=3, lwd=2)
lines(times, out_super$Iz/N, lty=2, lwd=2)
lines(times, out_super$Iwz/N, lty=4, lwd=2)

legend("topleft" , inset=.05, c("S Susceptible","Iw Sensitive strain", "Iz Resistant strain", "Iwz Double infected" ), lty=c(1,3,2,4), lwd=c(2,2,2,2))
```

__Plot n°2 Superinfection model__:$\beta$ ~w~ = `r pbeta_w`, $\beta$ ~z~ = `r pbeta_z`, $\gamma$ = `r pgammaa`, $\gamma$`= `r pgamma_tilt`, $\epsilon$ = `r peps` and $q$=`r pq`  are the chosen values of parameters. The $\delta$t = 1/12 year i.e. 1 month and we follow the evolution of the sensitive subpopulation (_solid line_), the subpopulation infected with sensitive strain exclusively (_dotted line_), with resistant strain exclusively (_dashed line_) and with both (_dashed-dotted line_) over `r a` years.


<center> <h2> Exclusive infection model </h2> </center>

This model is based on a relatively independent evolution of the resistant and sensitive strain which are in indirect competition, sharing hence only the susceptible population of hosts.

The model is defined by following equations:



\begin{aligned}

\frac{dS}{dt} &=-{\beta_w} \times{I_w}\times{S} -{\beta_z} \times{I_z}\times{S} + {\gamma} \times {I_z} + {\gamma}\times(1-{\epsilon})\times{I_w} + {\gamma `} \times {\epsilon} \times {I_w} \\

\frac{dI_w}{dt} &={\beta_w} \times{I_w}\times{S} - {\gamma}\times(1-{\epsilon})\times{I_w}- {\gamma `} \times {\epsilon} \times {I_w} \\

\frac{dI_z}{dt} &={\beta_z} \times{I_z}\times{S} - {\gamma} \times {I_z} \\

\end{aligned}








```{r exclusive_def, include=FALSE}
parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps)

state <- c(S=N-2, Iw=1, Iz=1)

times=seq(0, a, by=1/12)

SIS_exclusive=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*Iw*S - beta_z*Iz*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*Iw*S - gamma*(1-eps)*Iw - gamma_prim*eps*Iw
        dIz = beta_z*Iz*S - gamma*Iz
        list(c(dS, dIw, dIz))
       })
}

require(deSolve)

out_exc=as.data.frame(
  ode(y=state, times=times, func=SIS_exclusive, parms=parameters)
)

```




```{r exclusive_trial, echo=FALSE, fig.align='center'}
plot(times, out_exc$S/N, main="Plot n°3 - Exclusive infection model", xlab="time (yrs)", ylab="ratio", ylim=c(0,1), lty=1, lwd=2, type="l")
lines(times, out_exc$Iw/N, lty=3, lwd=2)
lines(times, out_exc$Iz/N, lty=2, lwd=2)

legend("topleft" , inset=.05, c("S Susceptible","Iw Sensitive strain", "Iz Resistant strain" ), lty=c(1,3,2), lwd=c(2,2,2))

```



__Plot n°3 Exclusive infection model__: $\beta$ ~w~ = `r pbeta_w`, $\beta$ ~z~ = `r pbeta_z`, $\gamma$ = `r pgammaa`, $\gamma$`= `r pgamma_tilt`, $\epsilon$ = `r peps` and $q$=`r pq`  are the chosen values of parameters. The $\delta$t = 1/12 year i.e. 1 month and we follow the evolution of  the sensitive subpopulation (_solid line_), the subpopulation infected with sensitive strain exclusively (_dotted line_) and with resistant strain exclusively (_dashed line_) over `r a` years.




<center> <h2> Replacement infection model </h2> </center>

In this model, when a patient infected with resistant strain is super-infected by sensistive strain, the latter will replace the first one.

The model is defined by following equations:



\begin{aligned}
\frac{dS}{dt} &=-{\beta_w} \times{I_w}\times{S} - {\beta_z} \times{I_z}\times{S} + {\gamma} \times {I_z} + {\gamma}\times(1-{\epsilon})\times{I_w} + {\gamma `} \times {\epsilon} \times {I_w} \\

\frac{dI_w}{dt} &={\beta_w} \times{I_w}\times{S} + ({\beta_w}-{\beta_z})\times{I_w}\times{I_z} - {\gamma}\times(1-{\epsilon})\times{I_w}- {\gamma `} \times {\epsilon} \times {I_w} \\

\frac{dI_z}{dt} &={\beta_z} \times{I_z}\times{S} - {\gamma} \times {I_z} - ({\beta_w} - {\beta_z})\times{I_w}\times{I_z} \\

\end{aligned}







```{r replacement_def, include=FALSE}
parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps)

state <- c(S=N-2, Iw=1, Iz=1)

times=seq(0, a, by=1/12)

SIS_replacement=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*Iw*S - beta_z*Iz*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*Iw*S - gamma*(1-eps)*Iw - gamma_prim*eps*Iw + (beta_w-beta_z)*Iw*Iz
        dIz = beta_z*Iz*S - gamma*Iz - (beta_w-beta_z)*Iw*Iz
        list(c(dS, dIw, dIz))
       })
}

require(deSolve)

out_rep=as.data.frame(
  ode(y=state, times=times, func=SIS_replacement, parms=parameters)
)

```




```{r replacement_trial, echo=FALSE, fig.align='center'}
plot(times, out_rep$S/N, main="Plot n°4 - Replacement infection model", xlab="time (yrs)", ylab="ratio", ylim=c(0,1), lty=1, lwd=2, type="l")
lines(times, out_rep$Iw/N, lty=3, lwd=2)
lines(times, out_rep$Iz/N, lty=2, lwd=2)

legend("topleft" , inset=.05, c("S Susceptible","Iw Sensitive strain", "Iz Resistant strain" ), lty=c(1,3,2), lwd=c(2,2,2))
```



__Plot n°4 Replacement infection model__: $\beta$ ~w~ = `r pbeta_w`, $\beta$ ~z~ = `r pbeta_z`, $\gamma$ = `r pgammaa`, $\gamma$`= `r pgamma_tilt` and $\epsilon$ = `r peps`  are the chosen values of parameters. The $\delta$t = 1/12 year i.e. 1 month and we follow the evolution of the sensitive subpopulation (_solid line_), the subpopulation infected with sensitive strain exclusively (_dotted line_) and with resistant strain exclusively (_dashed line_) over 30 years.

<center> <h2> Unidirectional conversion model </h2> </center>

In this model, a patient patient that is treated by antibiotics and that is infected by sensitive strain will become infected by a resistant one as the outcome of the antibiotic treatment. This step modelizes clinical treatment failure where instead of cleansing the infection, the pathogen becomes resistant. The parameter $\rho$ represents the probability or frequence of clinical failure.

The model is defined by following equations:



\begin{aligned}

\frac{dS}{dt} &=-{\beta_w} \times{I_w}\times{S} - {\beta_z} \times{I_z}\times{S} + {\gamma} \times {I_z} + {\gamma}\times(1-{\epsilon})\times{I_w} + {\gamma `} \times {\epsilon} \times {I_w} \\

\frac{dI_w}{dt} &={\beta_w}\times{I_w}\times{S}-{\gamma}\times(1-{\epsilon})\times{I_w}-{\gamma `}\times{\epsilon}\times{I_w}-{\rho}\times{\epsilon}\times{I_w} \\

\frac{dI_z}{dt} &={\beta_z} \times{I_z}\times{S} - {\gamma} \times {I_z} +{\rho}\times{\epsilon}\times{I_w} \\

\end{aligned}




```{r unidir_def, include=FALSE}
parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps, rho=prho )

state <- c(S=N-2, Iw=1, Iz=1)

times=seq(0, a, by=1/12)

SIS_unidir=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*Iw*S - beta_z*Iz*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*Iw*S - gamma*(1-eps)*Iw - gamma_prim*eps*Iw - rho*eps*Iw
        dIz = beta_z*Iz*S - gamma*Iz + rho*eps*Iw
        list(c(dS, dIw, dIz))
       })
}

require(deSolve)

out_uni=as.data.frame(
  ode(y=state, times=times, func=SIS_unidir, parms=parameters)
)

```




```{r unidir_trial, echo=FALSE, fig.align='center'}
plot(times, out_uni$S/N, main="Plot n°5 - Unidirectional conversion model", xlab="time (yrs)", ylab="ratio", ylim=c(0,1), lty=1, lwd=2, type="l")
lines(times, out_uni$Iw/N, lty=3, lwd=2)
lines(times, out_uni$Iz/N, lty=2, lwd=2)

legend("topleft" , inset=.05, c("S Susceptible","Iw Sensitive strain", "Iz Resistant strain" ), lty=c(1,3,2), lwd=c(2,2,2))
```



__Plot n°5 Unidirectional conversion model__: $\beta$ ~w~ = `r pbeta_w`, $\beta$ ~z~ = `r pbeta_z`, $\gamma$ = `r pgammaa`, $\gamma$`= `r pgamma_tilt`, $\epsilon$ = `r peps` and $\rho$ = `r prho` are the chosen values of parameters. The $\delta$t = 1/12 year i.e. 1 month and we follow the evolution of the sensitive subpopulation (_solid line_), the subpopulation infected with sensitive strain exclusively (_dotted line_) and with resistant strain exclusively (_dashed line_) over `r a` years.




<center> <h2> Bidirectional conversion model </h2> </center>

In addition to the preceding model, a transition from $Iz$ to $Iw$ is taken into account. In the absence of treatment, some resistant bacteria may revert into sensitive state. These will then be able of outcompeting the resistant one. $\phi$ is the rate at which this phenomenon takes place in untreated patients.

The model is defined by following equations:



\begin{aligned}
\frac{dS}{dt} &=-{\beta_w} \times{I_w}\times{S} - {\beta_z} \times{I_z}\times{S} + {\gamma} \times {I_z} + {\gamma}\times(1-{\epsilon})\times{I_w} + {\gamma `} \times {\epsilon} \times {I_w} \\

\frac{dI_w}{dt} &={\beta_w}\times{I_w}\times{S}-{\gamma}\times(1-{\epsilon})\times{I_w}-{\gamma `}\times{\epsilon}\times{I_w}-{\rho}\times{\epsilon}\times{I_w} \\

\frac{dI_z}{dt} &={\beta_z} \times{I_z}\times{S} - {\gamma} \times {I_z} +{\rho}\times{\epsilon}\times{I_w} \\


\end{aligned}




```{r bidir_def, include=FALSE}
parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps, rho=prho, phi=pphi )

state <- c(S=N-2, Iw=1, Iz=1)

times=seq(0, a, by=1/12)

SIS_bidir=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*Iw*S - beta_z*Iz*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*Iw*S - gamma*(1-eps)*Iw - gamma_prim*eps*Iw - rho*eps*Iw + phi*(1-eps)*Iz
        dIz = beta_z*Iz*S - gamma*Iz + rho*eps*Iw - phi*(1-eps)*Iz
        list(c(dS, dIw, dIz))
       })
}

require(deSolve)

out_bi=as.data.frame(
  ode(y=state, times=times, func=SIS_bidir, parms=parameters)
)

```




```{r bidir_trial, echo=FALSE, fig.align='center'}
plot(times, out_bi$S/N, main="Plot n°6", xlab="time (yrs)", ylab="ratio", ylim=c(0,1), lty=1, lwd=2, type="l")
lines(times, out_bi$Iw/N, lty=3, lwd=2)
lines(times, out_bi$Iz/N, lty=2, lwd=2)

legend("topleft" , inset=.05, c("S Susceptible","Iw Sensitive strain", "Iz Resistant strain" ), lty=c(1,3,2), lwd=c(2,2,2))
```



__Plot n°6 Bidirectional conversion model__: $\beta$ ~w~ = `r pbeta_w`, $\beta$ ~z~ = `r pbeta_z`, $\gamma$ = `r pgammaa`, $\gamma$`= `r pgamma_tilt`, $\epsilon$ = `r peps`, $\rho$ = `r prho` and $\phi$= `pphi` are the chosen values of parameters. The $\delta$t = 1/12 year i.e. 1 month and we follow the evolution of the sensitive subpopulation (_solid line_), the subpopulation infected with sensitive strain exclusively (_dotted line_) and with resistant strain exclusively (_dashed line_) over `r a`.



