---
title: "Steady-state exploration"
author: "Jakub"
date: "10 října 2016"
output: html_document
---





```{r include=FALSE}
library(deSolve)
options("scipen" = 10)
N <- 1
a <- 20000
pbeta_w <- 0.04
pbeta_z <- 0.015
pgammaa <- 0.01
pgamma_tilt <- 0.1
pq <- 0.5
prho <- 0.5
pphi <- 0.05
peps <- 0.15
m <- 1
prop <- 0.01
conv_crit=0.00001
pepsi <- seq(0, 0.90, by=0.001)
tab <- matrix(data=NA, nrow=length(pepsi), ncol=5)

```

<center> <h1> Steady-state exploration of Spicknall`s models </h1> </center>

The purpose of this part is to explore the steady states of models mentionned in Spicknall\`s review with $\epsilon$ varying from 0 to 0.9. For each model, the parameters and the initial conditions are the same, where applicable. Every model starts with S=`r N-2*prop`, I~w~=`r prop` and I~z~=`r prop` at t=0. We used th same parameters as in the basic models: $\beta$ ~w~ = `r pbeta_w`, $\beta$ ~z~ = `r pbeta_z`, $\gamma$ = `r pgammaa`, $\gamma$\`= `r pgamma_tilt`, $\rho$ = `r prho` and $\phi$= `r pphi`.

<h2> Single-strain model - steady states <h2>

``` {r sis_steady_state_eps, include=FALSE}

parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps)
state <- c(S1=1-prop, S2=1-prop, Iw=prop, Iz=prop)
times=seq(0, a, by=m)


SIS_single_strain=function(times, state, parameters){
  with(as.list(c(state,parameters)),
    {dS1 = - beta_w*Iw*S1 + gamma*(1-eps)*Iw + gamma_prim*eps*Iw
      dS2 = - beta_z*Iz*S2 + gamma*Iz
      dIw = beta_w*Iw*S1 - gamma*(1-eps)*Iw - gamma_prim*eps*Iw
      dIz = beta_z*Iz*S2 - gamma*Iz
      list(c(dS1, dS2, dIw, dIz))
      return(list(c(dS1, dS2, dIw, dIz)))
    })
}


root_sis=function(times, state, parameters){
  dstate <- unlist(SIS_single_strain(times, state, parameters))
  sum(abs(dstate)) - conv_crit}

  
  
exploit_param=function(x) {
  parameters[5] <- x
  out <- lsodar(func=SIS_single_strain, y=state, parms=parameters, times=times, rootfunc=root_sis)
  return(tail(out, n=1))
}

steady <- do.call(rbind, lapply(pepsi, exploit_param))
steady <- unlist(steady)

steady_time <- steady[,1]
steady_S1 <- steady[,2]
steady_S2 <- steady[,3]
steady_Iw <- steady[,4]
steady_Iz <- steady[,5]

```



```{r sis_steady_state_eps_plot, echo=FALSE, fig.align='center'}

plot(pepsi, steady_Iw, main="Single-strain model steady states", xlab="proportion of treated patients", ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(pepsi, steady_Iz, lty=2, lwd=2)

legend("topright" , inset=.05, c("Iw Sensitive strain", "Iz Resistant strain" ), lty=c(3,2), lwd=c(2,2))

```


<h2> Superinfection model - steady states <h2>

``` {r sup_steady_state_eps, include=FALSE}

parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps, q=pq )

state <- c(S=1-2*prop, Iw=prop, Iz=prop, Iwz=0)

times = seq(0, a, by=m)

SIS_sup=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*(Iw+q*Iwz)*S - beta_z*(Iz+q*Iwz)*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*(Iw+q*Iwz)*S + gamma*Iwz - beta_z*(Iz+q*Iwz)*Iw - gamma*(1-eps)*Iw - gamma_prim*eps*Iw 
        dIz = beta_z*(Iz+q*Iwz)*S + gamma*(1-eps)*Iwz + gamma_prim*eps*Iwz - beta_w*(Iw+q*Iwz)*Iz - gamma*Iz 
        dIwz = beta_w*(Iw+q*Iwz)*Iz + beta_z*(Iz+q*Iwz)*Iw - gamma*Iwz - gamma*(1-eps)*Iwz - gamma_prim*eps*Iwz
        list(c(dS, dIw, dIz, dIwz))
       })
  }

root_sup=function(times, state, parameters){
  dstate <- unlist(SIS_sup(times, state, parameters))
  sum(abs(dstate)) - conv_crit
  }

  
  
exploit_param_sup=function(x) {
  parameters[5] <- x
  out <- lsodar(func=SIS_sup, y=state, parms=parameters, times=times, rootfunc=root_sup)
  return(tail(out, n=1))
  }

steady_sup <- do.call(rbind, lapply(pepsi, exploit_param_sup))
steady_sup <- unlist(steady_sup)

steady_time_sup <- steady_sup[,1]
steady_S_sup <- steady_sup[,2]
steady_Iw_sup <- steady_sup[,3]
steady_Iz_sup <- steady_sup[,4]
steady_Iwz_sup <- steady_sup[,5]

```



```{r sup_steady_state_eps_plot, echo=FALSE, fig.align='center'}
plot(pepsi, steady_Iw_sup, main="Superinfection model steady states", xlab="proportion of treated patients", ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(pepsi, steady_Iz_sup, lty=2, lwd=2)
lines(pepsi, steady_S_sup, lty=1, lwd=2)
lines(pepsi, steady_Iwz_sup, lty=4, lwd=2)

legend("topright" , inset=.05, c("Iw Sensitive strain", "Iz Resistant strain", "Iwz Double Infected", "S Susceptible population" ), lty=c(3,2,4,1), lwd=c(2,2,2,2))

```

<h2> Exclusion model - steady states <h2>

``` {r exc_steady_state_eps, include=FALSE}

parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps)

state <- c(S=1-2*prop, Iw=prop, Iz=prop)

times=seq(0, a, by=m)

SIS_exc=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*Iw*S - beta_z*Iz*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*Iw*S - gamma*(1-eps)*Iw - gamma_prim*eps*Iw
        dIz = beta_z*Iz*S - gamma*Iz
        list(c(dS, dIw, dIz))
       })
}


root_exc=function(times, state, parameters){
  dstate <- unlist(SIS_exc(times, state, parameters))
  sum(abs(dstate)) - conv_crit
  }

  
  
exploit_param_exc=function(x) {
  parameters[5] <- x
  out <- lsodar(func=SIS_exc, y=state, parms=parameters, times=times, rootfunc=root_exc)
  return(tail(out, n=1))
  }

steady_exc <- do.call(rbind, lapply(pepsi, exploit_param_exc))
steady_exc <- unlist(steady_exc)

steady_time_exc <- steady_exc[,1]
steady_S_exc <- steady_exc[,2]
steady_Iw_exc <- steady_exc[,3]
steady_Iz_exc <- steady_exc[,4]

```


```{r exc_steady_state_eps_plot, echo=FALSE, fig.align='center'}
plot(pepsi, steady_Iw_exc, main="Exclusion model steady states", xlab="proportion of treated patients", ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(pepsi, steady_Iz_exc, lty=2, lwd=2)
lines(pepsi, steady_S_exc, lty=1, lwd=2)

legend("topright" , inset=.05, c("Iw Sensitive strain", "Iz Resistant strain", "S Susceptible population" ), lty=c(3,2,4), lwd=c(2,2,2))

```

<h2> Replacement model - steady states <h2>

``` {r rep_steady_state_eps, include=FALSE}

parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps)

state <- c(S=1-2*prop, Iw=prop, Iz=prop)

times=seq(0, a, by=m)

SIS_rep=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*Iw*S - beta_z*Iz*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*Iw*S - gamma*(1-eps)*Iw - gamma_prim*eps*Iw + (beta_w-beta_z)*Iw*Iz
        dIz = beta_z*Iz*S - gamma*Iz - (beta_w-beta_z)*Iw*Iz
        list(c(dS, dIw, dIz))
       })
}


root_rep=function(times, state, parameters){
  dstate <- unlist(SIS_rep(times, state, parameters))
  sum(abs(dstate)) - conv_crit
  }

  
  
exploit_param_rep=function(x) {
  parameters[5] <- x
  out <- lsodar(func=SIS_rep, y=state, parms=parameters, times=times, rootfunc=root_rep)
  return(tail(out, n=1))
  }

steady_rep <- do.call(rbind, lapply(pepsi, exploit_param_rep))
steady_rep <- unlist(steady_rep)

steady_time_rep <- steady_rep[,1]
steady_S_rep <- steady_rep[,2]
steady_Iw_rep <- steady_rep[,3]
steady_Iz_rep <- steady_rep[,4]

```

```{r rep_steady_state_eps_plot, echo=FALSE, fig.align='center'}
plot(pepsi, steady_Iw_rep, main="Replacement model steady states", xlab="proportion of treated patients", ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(pepsi, steady_Iz_rep, lty=2, lwd=2)
lines(pepsi, steady_S_rep, lty=1, lwd=2)

legend("topright" , inset=.05, c("Iw Sensitive strain", "Iz Resistant strain", "S Susceptible population" ), lty=c(3,2,4), lwd=c(2,2,2))

```

<h2> Unidirectional transition model - steady states <h2>

``` {r uni_steady_state_eps, include=FALSE}

parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps, rho=prho )

state <- c(S=1-2*prop, Iw=prop, Iz=prop)

times=seq(0, a, by=m)

SIS_uni=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*Iw*S - beta_z*Iz*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*Iw*S - gamma*(1-eps)*Iw - gamma_prim*eps*Iw - rho*eps*Iw
        dIz = beta_z*Iz*S - gamma*Iz + rho*eps*Iw
        list(c(dS, dIw, dIz))
       })
}


root_uni=function(times, state, parameters){
  dstate <- unlist(SIS_uni(times, state, parameters))
  sum(abs(dstate)) - conv_crit
  }

  
  
exploit_param_uni=function(x) {
  parameters[5] <- x
  out <- lsodar(func=SIS_uni, y=state, parms=parameters, times=times, rootfunc=root_uni)
  return(tail(out, n=1))
  }

steady_uni <- do.call(rbind, lapply(pepsi, exploit_param_uni))
steady_uni <- unlist(steady_uni)

steady_time_uni <- steady_uni[,1]
steady_S_uni <- steady_uni[,2]
steady_Iw_uni <- steady_uni[,3]
steady_Iz_uni <- steady_uni[,4]

```

```{r uni_steady_state_eps_plot, echo=FALSE, fig.align='center'}
plot(pepsi, steady_Iw_uni, main="Unidirectional transition model steady states", xlab="proportion of treated patients", ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(pepsi, steady_Iz_uni, lty=2, lwd=2)
lines(pepsi, steady_S_uni, lty=1, lwd=2)

legend("topright" , inset=.05, c("Iw Sensitive strain", "Iz Resistant strain", "S Susceptible population" ), lty=c(3,2,4), lwd=c(2,2,2))

```

<h2> Bidirectional transition model - steady states <h2>

``` {r bi_steady_state_eps, include=FALSE}

parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps, rho=prho, phi=pphi )

state <- c(S=1-2*prop, Iw=prop, Iz=prop)

times=seq(0, a, by=m)

SIS_bidir=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*Iw*S - beta_z*Iz*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*Iw*S - gamma*(1-eps)*Iw - gamma_prim*eps*Iw - rho*eps*Iw + phi*(1-eps)*Iz
        dIz = beta_z*Iz*S - gamma*Iz + rho*eps*Iw - phi*(1-eps)*Iz
        list(c(dS, dIw, dIz))
       })
}


root_bi=function(times, state, parameters){
  dstate <- unlist(SIS_bidir(times, state, parameters))
  sum(abs(dstate)) - conv_crit
  }

  
  
exploit_param_bi=function(x) {
  parameters[5] <- x
  out <- lsodar(func=SIS_bidir, y=state, parms=parameters, times=times, rootfunc=root_bi)
  return(tail(out, n=1))
  }

steady_bi <- do.call(rbind, lapply(pepsi, exploit_param_bi))
steady_bi <- unlist(steady_bi)

steady_time_bi <- steady_bi[,1]
steady_S_bi <- steady_bi[,2]
steady_Iw_bi <- steady_bi[,3]
steady_Iz_bi <- steady_bi[,4]

```

```{r bi_steady_state_eps_plot, echo=FALSE, fig.align='center'}
plot(pepsi, steady_Iw_bi, main="Bidirectional transition model steady states", xlab="proportion of treated patients", ylab="ratio", ylim=c(0,1), lty=3, lwd=2, type="l")

lines(pepsi, steady_Iz_bi, lty=2, lwd=2)
lines(pepsi, steady_S_bi, lty=1, lwd=2)

legend("topright" , inset=.05, c("Iw Sensitive strain", "Iz Resistant strain", "S Susceptible population" ), lty=c(3,2,1), lwd=c(2,2,2))

```