---
title: "Model_synthesis"
author: "Jakub"
date: "13 října 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center> <h1> Synthesis of models mentionned in Spicknall`s review </h1> </center>

<center> <h2> Single-strain model and Superinfection model </h2> </center>

In fact, the Single-strain model is but a special case of the Superinfection model where q=1. With q=1 there is no dampened infectiousness and each strain in a double-infected host will infect susceptible host at the same rate that single-infected host does. This situation is not too realistic for many as both strains usually compete for colonizing same sites and using same resources. Therefore this model with q=1 is relevant only in case that the resistant strain infects a different site than the susceptible one and these sites are independent.

The superinfection model is relevant only for pathogens where the resistant train and the sensitive one can coexist in high level. This is the case for _Streptococcus pneumoniae_, _Staphylococcus Aureus_ and _Mycobacterium tuberculosis_. 

\begin{aligned}

\frac{dS}{dt} &=-{\beta_w} \times{I_w}\times{S} -{\beta_z} \times{I_z}\times{S} + {\gamma} \times {I_z} + {\gamma}\times(1-{\epsilon})\times{I_w} + {\gamma `} \times {\epsilon} \times {I_w} \\

\frac{dI_w}{dt} &={\beta_w} \times{I_w}\times{S} - {\gamma}\times(1-{\epsilon})\times{I_w}- {\gamma `} \times {\epsilon} \times {I_w} \\

\frac{dI_z}{dt} &={\beta_z} \times{I_z}\times{S} - {\gamma} \times {I_z} \\

\end{aligned}








```{r sup1_def, include=FALSE, cache=TRUE}
parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps)

state <- c(S=1-2*prop, Iw=prop, Iz=prop)

times=seq(0, a, by=m)

SIS_sup1=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*Iw*S - beta_z*Iz*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*Iw*S - gamma*(1-eps)*Iw - gamma_prim*eps*Iw
        dIz = beta_z*Iz*S - gamma*Iz
        list(c(dS, dIw, dIz))
       })
}

require(deSolve)

out_sup1=as.data.frame(
  ode(y=state, times=times, func=SIS_sup1, parms=parameters)
)

```




```{r sup1_trial, echo=FALSE, fig.align='center'}
plot(times, out_sup1$S, main="Plot n°3 - Exclusive infection model", xlab="time (steps)", ylab="ratio", ylim=c(0,1), lty=1, lwd=2, type="l")
lines(times, out_sup1$Iw, lty=3, lwd=2)
lines(times, out_sup1$Iz, lty=2, lwd=2)

legend("topleft" , inset=.05, c("S Susceptible","Iw Sensitive strain", "Iz Resistant strain" ), lty=c(1,3,2), lwd=c(2,2,2))

```


\begin{aligned}
\frac{dS}{dt} &=-{\beta_w} \times{I_w}\times{S} - {\beta_z} \times{I_z}\times{S} + {\gamma} \times {I_z} + {\gamma}\times(1-{\epsilon})\times{I_w} + {\gamma `} \times {\epsilon} \times {I_w} \\

\frac{dI_w}{dt} &={\beta_w}\times{I_w}\times{S}-{\gamma}\times(1-{\epsilon})\times{I_w}-{\gamma `}\times{\epsilon}\times{I_w}-{\rho}\times{\epsilon}\times{I_w} \\

\frac{dI_z}{dt} &={\beta_z} \times{I_z}\times{S} - {\gamma} \times {I_z} +{\rho}\times{\epsilon}\times{I_w} \\


\end{aligned}




```{r sup2_def, include=FALSE, cache=TRUE}
parameters <- c(beta_w=pbeta_w, beta_z=pbeta_z, gamma=pgammaa, gamma_prim=pgamma_tilt, eps=peps, rho=prho, phi=pphi )

state <- c(S=1-2*prop, Iw=prop, Iz=prop)

times=seq(0, a, by=m)

SIS_sup2=function(times, state, parameters){
  with(as.list(c(state,parameters)),
       {dS = - beta_w*Iw*S - beta_z*Iz*S + gamma*(1-eps)*Iw + gamma_prim*eps*Iw + gamma*Iz
        dIw = beta_w*Iw*S - gamma*(1-eps)*Iw - gamma_prim*eps*Iw - rho*eps*Iw + phi*(1-eps)*Iz
        dIz = beta_z*Iz*S - gamma*Iz + rho*eps*Iw - phi*(1-eps)*Iz
        list(c(dS, dIw, dIz))
       })
}

require(deSolve)

out_sup2=as.data.frame(
  ode(y=state, times=times, func=SIS_sup2, parms=parameters)
)

```




```{r sup2_trial, echo=FALSE, fig.align='center'}
plot(times, out_sup2$S, main="Plot n°6", xlab="time (steps)", ylab="ratio", ylim=c(0,1), lty=1, lwd=2, type="l")
lines(times, out_sup2$Iw, lty=3, lwd=2)
lines(times, out_sup2$Iz, lty=2, lwd=2)

legend("topleft" , inset=.05, c("S Susceptible","Iw Sensitive strain", "Iz Resistant strain" ), lty=c(1,3,2), lwd=c(2,2,2))
```


